On exponential sums arising from the classical doubling method

### 2025-04-10 main.tex
Minors

We recall that $$
	\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n} = \prod_{a \in \mathbb{Z}/m\mathbb{Z}} \left(\algebraicGroup{\GL}_n\right)_a,
$$
with Frobenius root action $$\Frobenius(g_0,g_1, \hdots, g_{m-1}) = (\Frobenius(g_{m-1}), \Frobenius(g_0), \Frobenius(g_1), \hdots, \Frobenius(g_{m-2})).$$ We will freely identify $\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n}(\finiteField) = \GL_n(\finiteFieldExtension{m})$ via the map $\left(g_0,\dots,g_{m-1}\right) \mapsto g_0$.

@@

\begin{center}
	\begin{tabular}{|c|c|} \hline
		$\algebraicGroup{G}$ & $\algebraicGroup{G}^*$ representation  \tabularnewline \hline \hline
		$\algebraicGroup{GL}_n$ &  $\algebraicGroup{\GL}_n \to \algebraicGroup{GL}_{2n}$ \tabularnewline \hline
		$\algebraicGroup{\UnitaryGroup}_{n}$ &  $\algebraicGroup{\UnitaryGroup}_{n} \to \restrictionOfScalars{\quadraticExtension}{\finiteField}{\algebraicGroup{GL}_n}$\tabularnewline \hline		 				 
		$\algebraicGroup{\SO}_{2n+1}$ &$\algebraicGroup{\Sp}_{2n} \to \algebraicGroup{\GL}_{2n}$ \tabularnewline \hline
		$\algebraicGroup{\Sp}_{2n}$  &$\algebraicGroup{\SO}_{2n+1}\to \algebraicGroup{\GL}_{2n+1}$\tabularnewline \hline		 
		$\algebraicGroup{\SO}^{+}_{2n}$ &$\algebraicGroup{\SO}^{+}_{2n}\to \algebraicGroup{\GL}_{2n}$\tabularnewline \hline
		$\algebraicGroup{\SO}^{-}_{2n}$ &$\algebraicGroup{\SO}^{-}_{2n}\to \algebraicGroup{\GL}_{2n}$\tabularnewline \hline
		$\algebraicGroup{\GSp}_{2n}$  &$\algebraicGroup{GSpin}_{2n+1}\to \algebraicGroup{\GL}_{2n+1}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{+}_{2n}$ & $\algebraicGroup{GSpin}_{2n}^{+}\to\algebraicGroup{\GL}_{2n}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{-}_{2n}$& $\algebraicGroup{GSpin}_{2n}^{-}\to\algebraicGroup{\GL}_{2n}$\tabularnewline \hline
	\end{tabular}
\end{center}
Where in each case the map on cocharacters $\rho_*:X_*(\algebraicGroup{T}^*) \cong (\mathbb{Q}/\mathbb{Z})_{p'}^n \to X_*(\tilde{\algebraicGroup{T}}^*) \cong (\mathbb{Q}/\mathbb{Z})_{p'}^{2n}$ send $x$ to $(x,-w_n x)$ \calvin{in the GSpin cases might scale up by 2}
%Note that in the $SO$ and $GSO$ case, the map $\rho$ sends the split torus of $G^*$ into the split torus of $\GL(2n)$ and the induced map of cocharacters sends $x \mapsto (x,-x)$. In the $Sp,GSp$ cases the map is similar: the induced map of cocharacters sends $x \mapsto (x,1,-x)$. In the unitary case, the torus whose $\finiteField$ points are $(\NormOneGroup{2})^n \subset G^*(\finiteField)$ includes into the torus in $\GL(2n)$ whose $\finiteField$ points are $(\finiteFieldExtension{2}^\times)^n$ and the induced map of cocharacters is $(x_1,x_2, \hdots, x_n) \mapsto (x_1,-x_1,x_2,-x_2, \hdots ,x_n,-x_n)$.

%In particular 
%To analyze $g_{\algebraicGroup{T},\rho}(\chi,\theta,\psi)$ for these $\rho$, we need to understand what torus character pair of $\GL$ corresponds to $\phi((s))$ under a map $\phi(x) = (x,-x)$:
%The following computation in the proof of theorem \ref{lem:indep_geo_conj}.
Our proof of \Cref{lem:cochar_computation} will check that $\rho(s_{T,\theta}) = (s_{\Tilde{T},\Tilde{\theta}})$ with $g_{T}(\chi,\theta,\psi)$ equal to $\tau_{\Tilde{T}}(\Tilde{\theta} \times \chi,\psi)$ up to a few factors. We will do this by breaking our tori into irreducible rational factors with the following lemma handling each possible factor

>>>>

We recall that \elad{Maybe we can help the reader and tell them that $m=1,2$} $$
	\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n} = \prod_{a \in \mathbb{Z}/m\mathbb{Z}} \left(\algebraicGroup{\GL}_n\right)_a,
$$
with Frobenius root action $$\Frobenius(g_0,g_1, \hdots, g_{m-1}) = (\Frobenius(g_{m-1}), \Frobenius(g_0), \Frobenius(g_1), \hdots, \Frobenius(g_{m-2})).$$ We will freely identify $\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n}(\finiteField) = \GL_n(\finiteFieldExtension{m})$ via the map $\left(g_0,\dots,g_{m-1}\right) \mapsto g_0$.

@@

\begin{center}
	\begin{tabular}{|c|c|} \hline
		$\algebraicGroup{G}$ & $\rho \colon \algebraicGroup{G}^{\ast} \to \left(\restrictionOfScalars{\quadraticExtension}{\finiteField}{}\right) \algebraicGroup{\GL}_N$  \tabularnewline \hline \hline
		$\algebraicGroup{GL}_n$ &  $\algebraicGroup{\GL}_n \to \algebraicGroup{GL}_{2n}$ \tabularnewline \hline
		$\algebraicGroup{\UnitaryGroup}_{n}$ &  $\algebraicGroup{\UnitaryGroup}_{n} \to \restrictionOfScalars{\quadraticExtension}{\finiteField}{\algebraicGroup{GL}_n}$\tabularnewline \hline		 				 
		$\algebraicGroup{\SO}_{2n+1}$ &$\algebraicGroup{\Sp}_{2n} \to \algebraicGroup{\GL}_{2n}$ \tabularnewline \hline
		$\algebraicGroup{\Sp}_{2n}$  &$\algebraicGroup{\SO}_{2n+1}\to \algebraicGroup{\GL}_{2n+1}$\tabularnewline \hline		 
		$\algebraicGroup{\SO}^{+}_{2n}$ &$\algebraicGroup{\SO}^{+}_{2n}\to \algebraicGroup{\GL}_{2n}$\tabularnewline \hline
		$\algebraicGroup{\SO}^{-}_{2n}$ &$\algebraicGroup{\SO}^{-}_{2n}\to \algebraicGroup{\GL}_{2n}$\tabularnewline \hline
		$\algebraicGroup{\GSp}_{2n}$  &$\algebraicGroup{GSpin}_{2n+1}\to \algebraicGroup{\GL}_{2n+1}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{+}_{2n}$ & $\algebraicGroup{GSpin}_{2n}^{+}\to\algebraicGroup{\GL}_{2n}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{-}_{2n}$& $\algebraicGroup{GSpin}_{2n}^{-}\to\algebraicGroup{\GL}_{2n}$\tabularnewline \hline
	\end{tabular}
\end{center}
In each of the cases in the table, the map on cocharacters $\rho_{\ast}:\CocharacterLattice{\algebraicGroup{T}^{\ast}} \cong (\mathbb{Q}/\mathbb{Z})_{p'}^n \to \CocharacterLattice{\tilde{\algebraicGroup{T}}^{\ast}} \cong (\mathbb{Q}/\mathbb{Z})_{p'}^{2n}$ sends $x$ to $(x,-w_n x)$. \calvin{in the GSpin cases might scale up by 2}
%Note that in the $SO$ and $GSO$ case, the map $\rho$ sends the split torus of $G^*$ into the split torus of $\GL(2n)$ and the induced map of cocharacters sends $x \mapsto (x,-x)$. In the $Sp,GSp$ cases the map is similar: the induced map of cocharacters sends $x \mapsto (x,1,-x)$. In the unitary case, the torus whose $\finiteField$ points are $(\NormOneGroup{2})^n \subset G^*(\finiteField)$ includes into the torus in $\GL(2n)$ whose $\finiteField$ points are $(\finiteFieldExtension{2}^\times)^n$ and the induced map of cocharacters is $(x_1,x_2, \hdots, x_n) \mapsto (x_1,-x_1,x_2,-x_2, \hdots ,x_n,-x_n)$.

%In particular 
%To analyze $g_{\algebraicGroup{T},\rho}(\chi,\theta,\psi)$ for these $\rho$, we need to understand what torus character pair of $\GL$ corresponds to $\phi((s))$ under a map $\phi(x) = (x,-x)$:
%The following computation in the proof of theorem \ref{lem:indep_geo_conj}.
We will prove \Cref{lem:cochar_computation} by checking that if $\rho(s_{T,\theta}) = (s_{\Tilde{T},\Tilde{\theta}})$ then $g_{T}(\chi,\theta,\psi)$ equals to $\GaussSumTorusCharacter{\tilde{T}}{\tilde{\theta}}{\chi}{\fieldCharacter}$ up to some factors. We will do so by breaking our tori into irreducible rational factors with the following lemma handling each possible factor.
